# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r echo=TRUE}
data <- read.csv('activity.csv')
```

## What is mean total number of steps taken per day?
Group steps by date for which they were recorded:
```{r echo=TRUE}
per_day <- aggregate(steps ~ date, data, sum)
barplot(per_day$steps, main="Number of steps per day", names.arg=per_day$date)
```

In order to understand total number of steps a little bit more one could 
calculate the mean and median of the series:
```{r echo=TRUE}
mean_steps <- mean(per_day$steps)
median_steps <- median(per_day$steps)
```
Giving the following results:
- **Mean**: {r mean_steps}
- **Median**: {r median_steps} 

## What is the average daily activity pattern?

```{r echo=TRUE}
per_interval <- aggregate(steps ~ interval, data, mean)
x <- per_interval$interval
y <- per_interval$steps
plot(x, y, type="l")
```

One can clearly spot the maximum that appears at:
```{r echo=TRUE}
subset(per_interval, steps==max(y))$interval
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
